Empirically proving properties of an exponential distribution
========================================================

This document has been submitted as a part of Statistical Inference course.  
  
There are 4 parts to this report. Each part tries to answer one question asked as a part of the assignment.  

The following code will create the necessary data.  
```{r}
lambda <- 0.2
n <- 40
nosim <- 10000
expMeans <- vector()
expSD <- vector()
for (i in 1:nosim) {
    expMeans[i] <- mean(rexp(n, lambda))
    expSD[i] <- sd(rexp(n, lambda))
}
```

**Part 1**  
------------
Show where the distribution is centered at and compare it to the theoretical center of the distribution.  

The theoretical center for an exponential distribution is 1/lambda. In this case, it will be 1/0.2 = 5.  

Let's see what our data says.
```{r}
mean(expMeans)
```

Thus, both the mean and the median is very close to the theoretically center.  


**Part 2**  
------------
Show how variable it is and compare it to the theoretical variance of the distribution.  

The theoretical variance for our distribution will be  
(population variance)/(sample size) = (1/0.2)^2 /40 = 0.625  

Let's check the variance in our data
```{r}
var(expMeans)
```

Again, the theoretical variance and the computed variance are very close.  
With an increase in the number of simulations (variable 'nosim'), we will be able to see that the empirical evidence more strongly resembles the theoretically computed data.  


**Part 3**  
------------
Show that the distribution is approximately normal  

We know that approximately 68.27%, 95.45% and 99.73% of the normal density lies within 1, 2, and 3 standard deviations from the mean, respectively.  

Let's check this for our data.  
The code below gets the number of elements in expMeans which fall within 1 standard deviation of mean.  
```{r}
mn <- mean(expMeans)
stdDev <- sd(expMeans)
length(expMeans[expMeans > mn-stdDev & expMeans < mn+stdDev])
```
which is very close to 68.27% of 10000.

Similarly, we can check for 2 and 3 standard deviations from mean and see that they are approximately 95.45% and 99.73% of the values respectively.
```{r}
length(expMeans[expMeans > mn-2*stdDev & expMeans < mn+2*stdDev])
length(expMeans[expMeans > mn-3*stdDev & expMeans < mn+3*stdDev])
```

In addition, figure 1 in appendix compares a histogram of our data with a histogram created from an equal number of random normal deviates. The plots look similar.  

**Part 4**  
------------
Evaluate the coverage of the confidence interval for 1/lambda: $$X±1.96 * S/n^{1/2}$$  



APPENDIX
============ 
**Figure 1:** Comparing histogram of our data with a histogram created from an equal number of random normal deviates.  
```{r fig.width=6, fig.height=8}
rnormDist <- rnorm(10000, 5, 5/sqrt(40))

par(mfrow=c(2, 1))
hist(expMeans, breaks=100, xlim=c(3,7), main="Histogram of simulation data")
hist(rnormDist, breaks=100, xlim=c(3,7), main="Histogram of normal data")
```

